---
title: "No one took your job."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning= FALSE)
```




```{r include = FALSE}


library(plyr)
library(dplyr)
library(knitr)
#library(optimbase)
library(ggplot2)
library(data.table)

directory <- "/Users/emily.webber/Dropbox/Website Dropbox 2/PopulationData"
setwd(directory)
options(scipen = 999)

#FOREING BORN************************
FB <- read.csv("Foreign.csv")
P <- read.csv("Population.csv")
PV <- read.csv("UNEMP2.csv")

Total <- select(FB, Year, Total)
Native <- select(FB, Year, Native)
Foreign <- select(FB, Year, Foreign)
Foreign_NC <- select(FB, Year, Foreign_Not_Citizen)
Naturalized <- select(FB, Year, Naturalized)

Total$Group <- "Total Population"
Total$Color <- "blue"
colnames(Total)[colnames(Total)=="Total"] <- "Population"
Native$Group <- "Native Born Population"
Native$Color <- "green"
colnames(Native)[colnames(Native)=="Native"] <- "Population"
Foreign$Group <- "Foreign Born Population"
Foreign$Color <- "purple"
colnames(Foreign)[colnames(Foreign)=="Foreign"] <- "Population"
Foreign_NC$Group <- "Foreign Born Not Citizens"
Foreign_NC$Color <- "orange"
colnames(Foreign_NC)[colnames(Foreign_NC)=="Foreign_Not_Citizen"] <- "Population"
Naturalized$Group <- "Foreign Born Naturalized Citizens"
colnames(Naturalized)[colnames(Naturalized)=="Naturalized"] <- "Population"
Naturalized$Color <- "red"

Born_Graph <- rbind(Total, Native, Foreign, Foreign_NC, Naturalized)
Born_Graph$Population <- Born_Graph$Population/1000000



UNEMP <- read.csv("UNEMP2.csv")
UNEMP$Illegal <- UNEMP$Illegal * 1000000
cor.test(Foreign$Population, UNEMP$UN)
cor.test(Total$Population, UNEMP$UN)
cor.test(Foreign_NC$Population, UNEMP$UN)
cor.test(Native$Population, UNEMP$UN)
cor.test(Naturalized$Population, UNEMP$UN)


#*******************TABLE

US_Table1 <- NULL
US_Table1$Year <- Foreign$Year
US_Table1$Total <- Total$Population/1000000 
US_Table1$Native <- Native$Population/1000000  
US_Table1$Naturalized <- Naturalized$Population/1000000  
US_Table1$Foreign <- Foreign$Population/1000000 
US_Table1$'Foreign Not Citizens' <- Foreign_NC$Population/1000000  
US_Table1$Illegal <- UNEMP$Illegal/1000000 

US_Table1$Total <- round(US_Table1$Total, digits = 2)
US_Table1$Native <- round(US_Table1$Native, digits = 2)
US_Table1$Naturalized <- round(US_Table1$Naturalized, digits = 2)
US_Table1$Foreign <- round(US_Table1$Foreign, digits = 2)
US_Table1$'Foreign Not Citizens' <- round(US_Table1$'Foreign Not Citizens', digits = 2)
US_Table1$Illegal <- round(US_Table1$Illegal, digits = 2)


US_Table1 <- as.data.table(US_Table1)


kable(US_Table1)



#**********************POPULATION CHANGE STATES
library(choroplethr)
library(choroplethrMaps)
data(state.regions)


States_IMG <- select(P, REGION, NAME, INTERNATIONALMIG2010, INTERNATIONALMIG2011, INTERNATIONALMIG2012, INTERNATIONALMIG2013, INTERNATIONALMIG2014, INTERNATIONALMIG2015)

States_IMG <- filter(States_IMG, NAME != "United States" & NAME != "Northwest Region" &
                       NAME != "South Region" & NAME != "Midwest Region" &
                       NAME != "West Region" & NAME != "Northeast Region" & NAME != "Puerto Rico" )
         
States_IMG <- mutate(States_IMG, Increase = INTERNATIONALMIG2010 + INTERNATIONALMIG2011 + INTERNATIONALMIG2012 + INTERNATIONALMIG2013 + INTERNATIONALMIG2014 + INTERNATIONALMIG2015)

data(state.regions)
States_Map <- NULL
States_Map$region <- state.regions$region
States_Map$value <- States_IMG$Increase
States_Map <- as.data.frame(States_Map)
#States_Map$region2 <- state.abb
States_Map2 <- merge(States_Map, state.regions)

state_choropleth(States_Map2, num_colors = 7)

SUNE <- read.csv("SUNE.csv")
SUNE$region <- sapply(SUNE$region, tolower)

Crime <- read.csv("Crime.csv")
Crime$region <- sapply(Crime$region, tolower)

newdata <- States_IMG[order(States_IMG$INTERNATIONALMIG2015),] 

statescor <- merge(SUNE, States_Map, by = "region")
statescor2 <- merge(States_Map, Crime, by = "region")
statescor3 <- merge(SUNE, Crime, by = "region")


```


The purpose of this analysis was to examine potential relationships between the immigrant population size and the unemployment rate in the United States. 

Below is a graph that indicates the total native and immigrant population sizes in the united states from 2009-2015.   This data was obtained from publicly available [census records](https://factfinder.census.gov/faces/nav/jsf/pages/index.xhtml).   All data files and R code used in this analysis can be found on my [GitHub site](https://github.com/emilyswebber).

&nbsp;

##Population USA

Population size has gone up since 2009, and this is true for native born citizens, foreign born citizens and foreign born non-citizens. 

```{r fig.width= 8, fig.height = 4, echo = FALSE, warning = FALSE}

Born_Graph <- filter(Born_Graph, Group != 'Total Population' & Group != 'Foreign Born Population')

ggplot(Born_Graph, aes(Year, Population, fill=Group)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values=c("white","red", "blue")) +
  ggtitle("Population of USA") +
  labs(x = "Year", y = "Population (Millions)")

           
```

&nbsp;

##Population Breakdown USA (Millions)
```{r echo=FALSE}

kable(US_Table1)
```

&nbsp;

##Unemployment Rate

Despite the increase in population size, the unemployment rate has gone down since 2009.  This means that if there is a correlation between unemployment rate and immigration into the United States, that it would be a **negative** relationship.  This means that as immigration has gone **up**, that unemployment has gone **down**.  



```{r echo=FALSE}

ggplot(UNEMP, aes(Year, UN)) +
  geom_bar(stat = "identity", colour = "black", fill = "blue") +
  ggtitle("Unemployment Rate in USA") +
  labs(x = "Year", y = "Percentage")



```


Importantly, **correlations do not imply causation....EVER**!  What does this mean?  Just because numbers tend to go up and down together doesn't mean that one variable is **causing** the other to do anything.  More simply put, correlations could capture something that happens by coincidence.  This is true no matter how strong a correlation is.  Correlation analyses are good indicators of where to look deeper, and should never be interpreted as causal on their own. 
&nbsp;


##Correlation Analyses

First I ran a simple correlation on overall population size and unemployment rate.  This was statistically significant (*r* = -0.92, p = .002).  Results showed a strong inverse relationship between the total US population size and unemployment rate (*r* = 0.92) and this is unlikely due to chance (p = 0.002).
    
[What do the numbers in parentheses mean?](https://emilyswebber.github.io/PopulationData/about.html)

    
```{r echo=FALSE}

TotalPopulation <- Total$Population/1000000
plot(TotalPopulation, UNEMP$UN, main = "Total Population and Unemployment Rate",
     ylab = "Unemployment Rate", xlab = "Total Population (Millions)", col = "blue",
     ylim = c(0, 12))



```


```{r echo=FALSE}

#cor.test(Total$Population, UNEMP$UN)


```


There was also a negative (inverse) correlation between the unemployment rate and the foreign born population (*r* = -0.88, p = .009), but this relationship did not occur when I looked specifically at the **non citizen** foreign born population (*r* = -0.41, p = .36).  

```{r echo=FALSE}


ForeignPopulation <- Foreign$Population/1000000
plot(ForeignPopulation, UNEMP$UN, main = "Foreign Population and Unemployment Rate",
     ylab = "Unemployment Rate", xlab = "Foreign Population (Millions)", col = "blue",ylim = c(0, 12))

#cor.test(Foreign$Population, UNEMP$UN)


#Foreign_NCPopulation <- Foreign_NC$Population/1000000
#plot(Foreign_NCPopulation, UNEMP$UN, main = "Foreign Non-Citizen Population and Unemployment Rate",
     #ylab = "Unemployment Rate", xlab = "Foreign Non-Citizen Population (Millions)", col = "blue",
     #ylim = c(0, 12))

#cor.test(Foreign_NC$Population, UNEMP$UN)


```


Next, I did a correlation analysis on the native born and naturalized United States citizen population and unemployment rate.  The results showed statistically significant relationships between in both cases (Native: *r* = -0.94, p = .002; Naturalized: *r* = -0.92, p = .003).

```{r echo=FALSE}

NativePopulation <- Native$Population/1000000
plot(NativePopulation, UNEMP$UN, main = "Native Population and Unemployment Rate",
     ylab = "Unemployment Rate", xlab = "Native Population (Millions)", col = "blue",
     ylim = c(0, 12))




#cor.test(Native$Population, UNEMP$UN)

NaturalizedPopulation <- Naturalized$Population/1000000
plot(NaturalizedPopulation, UNEMP$UN, main = "Naturalized Population and Unemployment Rate",
     ylab = "Unemployment Rate", xlab = "Naturalized Population (Millions)", col = "blue",
     ylim = c(0, 12))


#cor.test(Naturalized$Population, UNEMP$UN)



```

Finally, I looked for a relationship between Unemployment Rate and the Illegal immigrant population size and did not find a significant relationship (*r* = -0.46, p = 0.30).  Census records do not ask for immigration status, this data was found [here](http://immigration.procon.org/view.resource.php?resourceID=000844).   


```{r echo=FALSE}

UNEMPIll <- UNEMP$Illegal/1000000
plot(UNEMPIll, UNEMP$UN, main = "Illegal Immigrant Population and Unemployment Rate",
     ylab = "Unemployment Rate", xlab = "Illegal Population (Millions)", col = "blue",
     ylim = c(0, 12))

#cor.test(UNEMP$Illegal, UNEMP$UN)

```

&nbsp;




##5 year total immigration by state

The United States is a big country, and I thought it might be worthwhile to illustrate where the most immigration is taking place.  To examine this I looked at immigration numbers for each state and totaled them for the past five years. 

Below is a map that indicates which states have had the heaviest influx of immigrants (both legal and illegal).  Again, this data was obtained via publicly available [census records on population change](http://www.census.gov/data/datasets/2016/demo/popest/nation-total.html). 

```{r echo=FALSE}


state_choropleth(States_Map2, num_colors = 7,
                 title = "Immigration by State")

```


I decided to dig a little deeper, and see if there is a correlation between unemployment and 5 year immigration numbers by state.  This gives us a much larger and more detailed data set than looking at the United States as a whole 

I found no relationship between [2016 state unemployment rates](https://www.bls.gov/web/laus/laumstrk.htm) and immigration numbers (*r* = 0.178, p = 0.21) or between [2014 violent crime rates](http://www.worldatlas.com/articles/the-most-dangerous-states-in-the-u-s.html) and immigration numbers (*r* = 0.226, p = 0.118).  However, when I examined the relationship between state unemployment rate and crime, there was a statistically significant correlation (*r* = 0.526, p < 0.001)

```{r echo=FALSE}

#cor.test(statescor$Unemployment, statescor$value)
#cor.test(statescor2$Crime, statescor2$value)
#cor.test(statescor3$Crime, statescor3$Unemployment)

plot(statescor3$Crime, statescor3$Unemployment, main = "Crime and Unemployment Rates",
     ylab = "Unemployment Rate (%)", xlab = "Crime (per 100K population)", col = "blue",
     ylim = c(0, 10))



```
&nbsp;

##Summary

1. All population sizes went up since 2009. 
&nbsp;

2. The unemployment rate fell from 9.9% to 6.3% (2009 - 2015).
&nbsp;

3. There were inverse correlations between population size and unemployment rate for the total US population, foreign born population, native born citizens and naturalized citizens.  There was not a significant relationship between population size of foreign born non-citizens and illegal immigrants. 
&nbsp;

4. When the data was broken down into a state-by-state analysis there was no relationship between immigration over the past 5 years and state unemployment rate.  Violent crime was also not correlated with changes in immigration.  However, unemployment and violent crime were significantly correlated when examined state-by-state. 
&nbsp;

##Discussion Questions

Data should be used to both raise and answer questions.  The analysis provided here answered some basic questions about recent immigration and unemployment rate in the United States, but raises many more.  

I have vowed to keep my personal opinion off of this website, so I will refrain from sharing it with you here.  Instead I will offer some discussion questions to get those neurons firing. 
&nbsp;


1. What are the limitations of this analysis?  There are always, always, always limitations to any analysis.  One example here: the data covered here only go back to 2009, the scope of this analysis is limited. 
&nbsp;

2. Unemployment rate overall seems to be going down as population size is going up, but we know correlation does not imply causation.  What other factors could be at play? 
&nbsp;

3. If the population size of foreign born non-citizens is not related to unemployment rate in the United States, in what other ways may this be impacting our country? 


<!--html_preserve-->

<link rel="stylesheet" href="style.css" type="text/css" />
<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-91308049-1', 'auto');
  ga('send', 'pageview');

</script>



<!-- Go to www.addthis.com/dashboard to customize your tools --> <script type="text/javascript" src="//s7.addthis.com/js/300/addthis_widget.js#pubid=ra-589036a8549be1ce"></script>


<!--/html_preserve-->
